Viewing Computer Science through Citation Analysis

### 2019-12-20 Scientometrics/comp.tex
update

\emph{Co-Citation.} For an alternate view of these DBLP data, we constructed clusters using co-citation, the frequency with which a pair of articles is cited by other articles \cite{small_co-citation_1973,marshakova-shaikevich_co-citation_1973}. Co-citation, first described independently by Small and Marshakova in 1973 \cite{marshakova-shaikevich_co-citation_1973}, provides insight into the emergence of new ideas derived from the association of previously independent ones. Unlike clustering by direct citation, where every input publication is assigned to a cluster and every citation is weighted equally, the co-citation relationship between papers is weighted to represent the strength of the co-citation history.
Because this produces a weighted graph, clustering methods that address weights are required.
Furthermore, co-citation produces relatively weak inter-cluster interactions, requiring additional modifications to standard clustering approaches \cite{boyack_cocitation_2010,boyack_improving_2013,small_structure_1974,small_clustering_1985}.

@@

Figure~\ref{fig:heatmap} permits examination of these data from two perspectives: (i) rows: the clusters that map to a given ASJC minor subject area and (ii) columns: ASJC minor subject areas that comprised at least 15\% of the publication in a cluster. Under these conditions of clustering and this threshold of 15\%, 31 of the 334 ASJC minor subject areas are represented. Unsurprisingly, the broad categories Computer Science Applications, Software, and Electrical and Electronic Engineering register in 16, 12, and 10 clusters respectively, while Artificial Intelligence mapped to 7 different clusters. At the other end of the range, 12 of the 31 ASJC minor subject areas were each detected in a single cluster.

@@

At the threshold of 15\%, only 17 of 20 co-citation clusters mapped to at least one minor subject areas. In contrast to a maximum of eight minor subject areas for clustering by direct citation at the 15\% threshold (Fig. 3, Cluster 17), no co-citation cluster maps to more than three minor subject areas. The threshold had to be reduced to 10\% for all 20 co-citation clusters to map to at least one minor subject area. We interpret these data as indicative of broader clusters created by weaker linkages that accumulate during the agglomerative clustering phase~\cite{small_clustering_1985}.

@@

Reconciling direct citation clusters to the Scopus ASJC classification yielded partially overlapping results that are consistent with the observation of Waltman and van Eck (2012)~\cite{waltman_new_2012} `that traditional disciplines such as those just mentioned only partly reflect the actual organization of today’s scientific research'. Of interest to us was the single obviously multidisciplinary cluster in which at least 15\% of its component publications were labeled with the ASJC minor subject areas Biochemistry, Chemistry, Computational Mathematics, Computational Theory and Mathematics, Computer Science Applications, Genetics, Molecular Biology, Statistics \& Probability. A second cluster mapped to Artificial Intelligence, Cognitive Neuroscience, and Neurology.  Both suggest collaboration between computer science and biology. At the opposite end of this spectrum is the perhaps more insular cluster that maps to Hardware and Architecture, Electrical and Electronic Engineering, and Software. These data suggest that, at least from the perspective of a high level of aggregation, some subfields within computer science may be primarily inward looking (remain concerned with fundamental questions in computer science and electrical engineering), while others are more actively engaged with fields external to computer science.

@@

In the 40 years since Salton and Bergmark's landmark paper, the field of computer science has not only expanded in volume, it has expanded in its interactions with other fields, and has
also resulted in new disciplinary and interdisciplinary subfields. 
Furthermore, machine learning and data science, which build off computer science and statistics, 
are emerging as major fields that are driving innovation in industry and science, and research in these areas is increasingly being performed in
fields external to computer science.
While DBLP provides an insight into what is commonly accepted as computer science, additional evaluation of the broader literature that
uses and develops computer science is needed to better assess the impact of computer science.

>>>>

\emph{Co-Citation.} For an alternate view of these DBLP data, we constructed clusters using co-citation, the frequency with which a pair of articles is cited by other articles \cite{small_co-citation_1973,marshakova-shaikevich_co-citation_1973}. Co-citation, first described independently by Small and Marshakova in 1973 \cite{marshakova-shaikevich_co-citation_1973}, provides insight into the emergence of new ideas derived from the association of previously independent ones. Unlike clustering by direct citation, where every input publication is assigned to a cluster and every citation is weighted equally, the co-citation relationship between papers is weighted to represent the strength of the co-citation history. Because this produces a weighted graph, clustering methods that address weights are required. Clustering by co-citation also considers weak inter-cluster interactions that involve modifications to standard clustering approaches. \cite{boyack_cocitation_2010,boyack_improving_2013,small_structure_1974,small_clustering_1985}.

@@

Figure~\ref{fig:heatmap} permits examination of these data from two perspectives: (i) rows: the clusters that map to a given ASJC minor subject area and (ii) columns: ASJC minor subject areas that comprised at least 15\% of the publication in a cluster. Under these conditions of clustering and this threshold of 15\%, 31 of the 334 ASJC minor subject areas are represented. Unsurprisingly, the broad categories Computer Science Applications, Software, and Electrical and Electronic Engineering register in 16, 12, and 10 clusters respectively, while Artificial Intelligence mapped to 7 different clusters. At the other end of the range, 12 of the 31 ASJC minor subject areas were each detected only in a single cluster.

@@

At the threshold of 15\%, only 17 of 20 co-citation clusters mapped to at least one minor subject areas. At the 15\% threshold, no co-citation cluster maps to more than three minor subject areas, in contrast to a maximum of eight minor subject areas for clustering by direct citation (Fig. 3, Cluster 17), . The threshold had to be reduced to 10\% for all 20 co-citation clusters to map to at least one minor subject area. We interpret these data as indicative of broader clusters created by weaker linkages that accumulate during the agglomerative clustering phase~\cite{small_clustering_1985}.

@@

Reconciling direct citation clusters to the Scopus ASJC classification yielded partially overlapping results that are consistent with the observation of Waltman and van Eck (2012)~\cite{waltman_new_2012} `that traditional disciplines such as those just mentioned only partly reflect the actual organization of today’s scientific research'. Of interest to us was the single obviously multidisciplinary cluster in which at least 15\% of its component publications were labeled with the ASJC minor subject areas Biochemistry, Chemistry, Computational Mathematics, Computational Theory and Mathematics, Computer Science Applications, Genetics, Molecular Biology, Statistics \& Probability. A second cluster mapped to Artificial Intelligence, Cognitive Neuroscience, and Neurology.  Both suggest collaboration between computer science and biology. At the opposite end of this spectrum is the cluster that maps to Hardware and Architecture, Electrical and Electronic Engineering, and Software. These data suggest that, at least from the perspective of a high level of aggregation, some subfields within computer science may be primarily inward looking (remain concerned with fundamental questions in computer science and electrical engineering), while others are more actively engaged with fields external to computer science.

@@

In the 40 years since Salton and Bergmark's landmark paper, the field of computer science has not only expanded in volume, it has expanded in its interactions with other fields, and has also resulted in new disciplinary and interdisciplinary subfields. Furthermore, machine learning and data science, which build off computer science and statistics, are emerging as major fields that are driving innovation in industry and science, and research in these areas is increasingly being performed in fields external to computer science.While DBLP provides an insight into what is commonly accepted as computer science, additional evaluation of the broader literature that
uses and develops computer science is needed to better assess the impact of computer science.
